%% -*- coding:utf-8 -*- 
\input preamble.tex

\begin{document}
\Russian
\input title.tex

\subsection*{Ticket 1} 
\begin{enumerate}
\item Decomposition of the electromagnetic field into modes (types of oscillations).
Hamiltonian form of the electromagnetic field equations.
Quantization of the electromagnetic field. 
\item Emission and absorption of light by an atom.
Hamiltonian of the atom-field system.
\end{enumerate}

\subsection*{Ticket 2} 
\begin{enumerate}
\item Properties of the operators $ \hat a $ and $ \hat a ^\dag $ 
Quantum state of the electromagnetic field with a defined
  energy. 
\item Interaction of the electromagnetic field of a resonator
(harmonic oscillator) with a reservoir of atoms at
temperature $T$.
Equation for the density matrix of the field in the number
representation.
\end{enumerate}

\subsection*{Ticket 3} 
\begin{enumerate}
\item Coherent states. 
\item Interaction of an atom with a mode of the electromagnetic field.
Relaxation of the dynamical system. Density matrix method.
\end{enumerate}

\subsection*{Ticket 4} 
\begin{enumerate}
\item Mixed states of the electromagnetic field.
Representation of the density operator using coherent
states. 
\item Laser model. Theory of laser generation. Statistics of laser
photons. 
\end{enumerate}

\subsection*{Ticket 5} 
\begin{enumerate}
\item Mixed states of the electromagnetic field.
Representation of the density operator using coherent
states. 
\item Polarization properties of light. Stokes parameters. EPR paradox
  for the Stokes parameters and entangled states.  
\end{enumerate}

\subsection*{Ticket 6} 
\begin{enumerate}
\item Decomposition of the electromagnetic field into modes (types of oscillations).
Hamiltonian form of the electromagnetic field equations.
Quantization of the electromagnetic field. 
\item Cryptography. Problems of classical cryptography. Quantum
  cryptography. 
\end{enumerate}

\end{document}