%% -*- coding:utf-8 -*- 
\documentclass[12pt,a4paper]{article} 
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}
\usepackage[dvips]{color}
\usepackage[xdvi]{graphics}
\usepackage{longtable}

\begin{document}
\Russian

\title{Examination Questions \\ ``Quantum Optics''}
%\author{I.\,V.~Murashko\thanks{e-mail: bourbaki@mail.ru}}
\author{}
\date{Saint Petersburg, 2015}
\maketitle
\newpage


\section*{1st Semester}

\subsection*{Question 1} 
\begin{enumerate}
\item Decomposition of the electromagnetic field into modes (types of oscillations).
Hamiltonian form of the electromagnetic field equations. Quantization
of the electromagnetic field. 
\item Interaction of the resonator electromagnetic field
  (harmonic oscillator) with a reservoir of atoms at temperature $T$. Equation for the field density matrix in the number representation.
\end{enumerate}

\subsection*{Question 2} 
\begin{enumerate}
\item Decomposition of the field into plane waves in free space. 
Density of states. Hamiltonian form of the field equations under decomposition into plane waves. Quantization of the electromagnetic field when decomposed into plane waves.
\item Relaxation of a dynamical system. Density matrix method. 
\end{enumerate}

\subsection*{Question 3} 
\begin{enumerate}
\item Properties of the operators $ \hat a $ and $ \hat a ^\dag $. Quantum
state of the electromagnetic field with a definite energy. 
\item Interaction of the resonator electromagnetic field
  (harmonic oscillator) with a reservoir of atoms at temperature $T$. Equation of motion for the statistical operator of the field mode in the coherent state representation.
\end{enumerate}

\subsection*{Question 4} 
\begin{enumerate}
\item Multimode states. 
\item Emission and absorption of light by an atom. 
Hamiltonian of the atom-field system.
\end{enumerate}

\subsection*{Question 5} 
\begin{enumerate}
\item Coherent states. 
\item Spontaneous emission. Weisskopf-Wigner approximation.
\end{enumerate}

\subsection*{Question 6} 
\begin{enumerate}
\item Mixed states of the electromagnetic field. 
\item Interaction
of an atom with a mode of the electromagnetic field. 
\end{enumerate}

\subsection*{Question 7} 
\begin{enumerate}
\item Representation of the density operator through coherent
  states.
\item Interaction of an atom with a multimode field. Spontaneous transitions.
\end{enumerate}

\section*{2nd Semester}

\subsection*{Question 1} 
\begin{enumerate}
\item Laser model
\item Coherent properties of light.
 First-order coherence.
\end{enumerate}

\subsection*{Question 2} 
\begin{enumerate}
\item Theory of laser generation
\item Squeezed states: applications of squeezed states. 
\end{enumerate}

\subsection*{Question 3} 
\begin{enumerate}
\item Statistics of laser photons
\item Coherent properties of light.
 Second-order coherence.
\end{enumerate}

\subsection*{Question 4} 
\begin{enumerate}
\item Laser theory. Coherent state representation
\item Photoelectric effect
\end{enumerate}

\subsection*{Question 5} 
\begin{enumerate}
\item Statistics of laser photons
\item Laser theory. Coherent state representation. Natural
  linewidth of laser emission.
\end{enumerate}

\subsection*{Question 6} 
\begin{enumerate}
\item Laser model
\item Non-classical light.
\end{enumerate}

\subsection*{Question 7} 
\begin{enumerate}
\item Theory of laser generation
\item Photon counting experiments. Applications of photon counting techniques to spectral measurements.
\end{enumerate}

\subsection*{Question 8} 
\begin{enumerate}
\item Quantum description of optical interference experiments
\item Non-classical light.
\end{enumerate}

\subsection*{Question 9} 
\begin{enumerate}
\item Mach–Zehnder interferometer
\item Squeezed states. Squeezing of quadrature states. Generation
  of squeezed states. Observation of squeezed states
\end{enumerate}

\subsection*{Question 10} 
\begin{enumerate}
\item Laser theory. Coherent state representation. Natural
  linewidth of laser emission.
\item Entangled states: definition, generation, detection
\end{enumerate}

\subsection*{Question 11} 
\begin{enumerate}
\item Coherent properties of light. First-order coherence
\item Entangled states: applications. Quantum
  teleportation. Quantum cryptography
\end{enumerate}

\subsection*{Question 12} 
\begin{enumerate}
\item Coherent properties of light. Second-order coherence
\item Bell inequalities. Non-classicality of entangled states
\end{enumerate}

\end{document}