%% -*- coding:utf-8 -*- 
\documentclass[12pt,a4paper]{article} 
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage[dvips]{color}
\usepackage[xdvi]{graphics}
\usepackage{longtable}

\begin{document}
\Russian

\title{Examination Questions \\ ``Quantum Optics''}
\author{I.\,V.~Murashko\thanks{e-mail: ivan.murashko@gmail.com}}
\date{Saint Petersburg, 2018-2019}
\maketitle
\newpage


\section*{1st Semester}

\subsection*{Question 1} 
\begin{enumerate}
\item Expansion of the electromagnetic field in modes (types of oscillations).
Hamiltonian form of the electromagnetic field equations. Quantization of the
electromagnetic field.
\item Interaction of an atom with a mode of the electromagnetic field.
\end{enumerate}

\subsection*{Question 2} 
\begin{enumerate}
\item Expansion of the field in plane waves in free space.
Density of states. Hamiltonian form of the field equations when expanded in plane
waves. Quantization of the electromagnetic field using its expansion in
plane waves.
\item Coherent states.
\end{enumerate}

\subsection*{Question 3} 
\begin{enumerate}
\item Properties of operators $ \hat a $ and $ \hat a ^\dag $. Quantum
state of the electromagnetic field with a definite energy.
\item Quantum nondemolition measurements. Ramsey interferometer.
\end{enumerate}

\subsection*{Question 4} 
\begin{enumerate}
\item Multimode states.
\item Emission and absorption of light by an atom.
Hamiltonian of the atom-field system.
\end{enumerate}

\subsection*{Question 5} 
\begin{enumerate}
\item Coherent states.
\item Spontaneous emission. Weisskopf-Wigner approximation.
\end{enumerate}

\subsection*{Question 6} 
\begin{enumerate}
\item Mixed states of the electromagnetic field.
\item Interaction of
an atom with a mode of the electromagnetic field.
\end{enumerate}

\subsection*{Question 7} 
\begin{enumerate}
\item Representation of the density operator via coherent
  states.
\item Spontaneous emission. Weisskopf-Wigner approximation.
\end{enumerate}

\subsection*{Question 8} 
\begin{enumerate}
\item Properties of creation operators $\hat{a}^\dag$ and annihilation operators $\hat{a}$.
\item Ramsey interferometer. Quantum nondemolition measurements.
\end{enumerate}

\section*{2nd Semester}


\subsection*{Question 1} 
\begin{enumerate}
\item Laser model
\item Coherent properties of light.
First-order coherence.
\end{enumerate}

\subsection*{Question 2} 
\begin{enumerate}
\item Theory of laser generation
\item Squeezed states: applications of squeezed states.
\end{enumerate}

\subsection*{Question 3} 
\begin{enumerate}
\item Statistics of laser photons
\item Coherent properties of light.
Second-order coherence.
\end{enumerate}

\subsection*{Question 4} 
\begin{enumerate}
\item Laser theory. Representation of coherent states
\item Photoelectric effect
\end{enumerate}

\subsection*{Question 5} 
\begin{enumerate}
\item Statistics of laser photons
\item Equation for the field density matrix in the number representation
\end{enumerate}

\subsection*{Question 6} 
\begin{enumerate}
\item Laser model
\item Nonclassical light.
\end{enumerate}

\subsection*{Question 7} 
\begin{enumerate}
\item Theory of laser generation
\item Bell inequalities. Nonclassicality of entangled states
\end{enumerate}

\subsection*{Question 8} 
\begin{enumerate}
\item Quantum description of optical interference experiments
\item Nonclassical light.
\end{enumerate}

\subsection*{Question 9} 
\begin{enumerate}
\item Mach-Zehnder interferometer
\item Squeezed states. Quadrature squeezing. Generation
  of squeezed states. Observation of squeezed states.
\end{enumerate}

\subsection*{Question 10} 
\begin{enumerate}
\item Laser theory. Representation of coherent states. Natural
  linewidth of laser radiation.
\item Entangled states: definition, generation, detection
\end{enumerate}

\subsection*{Question 11} 
\begin{enumerate}
\item Coherent properties of light. First-order coherence
\item Entangled states: applications. Quantum
  teleportation.
\end{enumerate}

\subsection*{Question 12} 
\begin{enumerate}
\item Coherent properties of light. Second-order coherence
\item Bell inequalities. Nonclassicality of entangled states
\end{enumerate}

\subsection*{Question 13} 
\begin{enumerate}
\item Interaction of the electromagnetic field of a resonator
  (harmonic oscillator) with a reservoir of atoms at
  temperature $T$.
\item Photoelectric effect
\end{enumerate}

\end{document}
