\chapter{Quantum Theory of the Laser}
\label{chLaser}

The semiclassical theory of the laser cannot answer all the questions that arise in connection with its operation. According to this theory, the laser does not generate at all before reaching the threshold, and when the threshold is exceeded, it begins to generate a classical electromagnetic field (light).

In reality, noticeably below the threshold, the laser generates chaotic light, and significantly above the threshold, its emission is close to classical. At and near the threshold, there is a transition region from chaotic light to ordered emission. Only a fully quantum theory can adequately describe this.

Another task, also requiring quantization of the electromagnetic field, is determining the ultimate (natural) linewidth of laser emission when the linewidth is determined by quantum fluctuations of the field, and various external influences, fundamentally eliminable, are not considered.

The laser is an open system in which active atoms and the resonator field are connected to large external systems that we will call reservoirs, providing pumping and losses. Therefore, the laser, as an open system, must be considered using a density matrix.

\input ./part2/laser/model.tex
\input ./part2/laser/theory.tex
\input ./part2/laser/stat.tex
\input ./part2/laser/theorycoh.tex
\input ./part2/laser/questions.tex

%% \begin{thebibliography}{99}
%% \bibitem{bCh1LaserMandel} L. Mandel, E. Wolf. Optical Coherence and
%%   Quantum Optics. Translated from English/Edited by V.V. Samartsev - M.:
%%   Nauka. Fizmatlit, 2000.- 896p.
%% \bibitem{bCh1Laser_Lamb} Lectures by W. Lamb//Quantum Optics and Quantum
%%   Radiophysics: Lectures at Summer School. M.: Mir, 1966.
%% \bibitem{bCh1Laser_Scally} Scully, M. Quantum Theory of the Laser -
%%   Problem of Nonequilibrium Statistical Mechanics. Quantum Fluctuations
%%   of Laser Emission/Compiled by F. Arecki, M. Scully, G. Haken, W. Weidlich. M.:
%%   Mir, 1974.
%% \bibitem{bCh1Laser_Haken} G. Haken. Laser Light Dynamics. M.: Mir,
%%   1988.
%% \bibitem{bCh1LaserSkalliZubari} M.O. Scully, M.S. Zubairy. Quantum
%%   Optics. M. Fizmatlit, 2003.
%% \end{thebibliography}