%% -*- coding:utf-8 -*- 
\section{Relaxation of a dynamical system. Density matrix method}
In quantum electronics and quantum optics, we deal with open
systems. Such systems must be considered using
the density matrix. We consider a small dynamical system
(atom, molecules, resonator mode, etc.) weakly coupled to
a large system called a reservoir (dissipative
system, thermostat), consisting of a large number of atoms,
molecules or oscillators, in equilibrium at
a certain temperature. Thus, in the general case,
we consider a dynamical system ($A$) interacting with
a reservoir ($B$) (see \autoref{figPart1Ch2_add2}). The density operator
of the combined system ($AB$) is denoted by $\hat{\rho}_{AB}$.
The density operator traced over the variables of the reservoir (thermostat) 
describes the behavior of the dynamical system taking into account the averaged
interaction with the thermostat. We have:
\begin{equation}
\hat{\rho}_{A} = Sp_{B} \hat{\rho}_{AB} = 
\sum_{\left(B\right)}
\bra{B}\hat{\rho}_{AB}\ket{B}.
\end{equation}
At the initial time $t_0$, the dynamical system and the reservoir
are considered independent (uncorrelated). Then
\begin{equation}
\hat{\rho}_{AB}\left(t_0\right) = 
\hat{\rho}_{A}\left(t_0\right)
\otimes
\hat{\rho}_{B}\left(t_0\right)
\end{equation}
where $Sp_B \hat{\rho}_{B} = 1$, $\otimes$ denotes the external or
tensor product. \rindex{Tensor product} 

\input ./part1/interaction/fig_add2.tex

The equation of motion for the density operator in the interaction picture
is
\[
\frac{d}{d t}\hat{\rho}_{AB} = 
- \frac{i}{\hbar}\left[\hat{V}_{AB}, \hat{\rho}_{AB}\right],
\]
where $\hat{V}_{AB}$ is the interaction Hamiltonian between systems $A$ 
and $B$. This equation can be solved using the following iterative process.
\begin{equation}
\hat{\rho}_{AB}\left(t\right) = 
\hat{\rho}^{(0)}_{AB}\left(t\right) +
\hat{\rho}^{(1)}_{AB}\left(t\right) +
\hat{\rho}^{(2)}_{AB}\left(t\right) + \dots,
\label{eqCh2_rho_iter}
\end{equation}
where the term \(\hat{\rho}^{(0)}_{AB}\left(t\right)\) is found under the assumption
of no interaction, i.e.
\[
\hat{\rho}^{(0)}_{AB}\left(t\right) = \hat{\rho}_{AB}\left(t_0\right) = 
\hat{\rho}_{A}\left(t_0\right)
\otimes
\hat{\rho}_{B}\left(t_0\right).
\]
The subsequent terms of the iterative relation \eqref{eqCh2_rho_iter} are given by
\begin{eqnarray}
\hat{\rho}^{(s)}_{AB}\left(t\right) = - \frac{i}{\hbar}\int_{t_0}^{t} dt_s
\left[\hat{V}_{AB}\left(t_s\right), \hat{\rho}^{(s - 1)}_{AB}\left(t_s\right)
\right] = 
\nonumber \\
= \left(- \frac{i}{\hbar}\right)^{s}\int_{t_0}^{t} dt_{s}
\int_{t_0}^{t_s} dt_{s -1} \dots
\nonumber \\
\dots \int_{t_0}^{t_2} dt_1 
\left[\hat{V}_{AB}\left(t_s\right), 
\left[\hat{V}_{AB}\left(t_{s -1}\right),
\dotsc
\left[\hat{V}_{AB}\left(t_1\right),
\hat{\rho}^{(0)}_{AB}\left(t_0\right)
\right]
\dots
\right]
\right] =
\nonumber \\
= \left(- \frac{i}{\hbar}\right)^{s}\int_{t_0}^{t} dt_{s}
\int_{t_0}^{t_{s}} dt_{s - 1} \dots
\nonumber \\
\dots \int_{t_0}^{t_2} dt_1 
\left[\hat{V}_{AB}\left(t_s\right), 
\left[\hat{V}_{AB}\left(t_{s-1}\right),
\dotsc
\left[\hat{V}_{AB}\left(t_1\right),
\hat{\rho}_{A}\left(t_0\right)
\otimes
\hat{\rho}_{B}\left(t_0\right)
\right]
\dotsc
\right]
\right].
\label{eqCh2_rho_iter_member}
\end{eqnarray}

Substituting \eqref{eqCh2_rho_iter_member} into \eqref{eqCh2_rho_iter} and performing averaging 
over the thermostat variables, we obtain
\begin{eqnarray}
\hat{\rho}_{A}\left(t\right) = 
\hat{\rho}_{A}\left(t_0\right) + 
\nonumber \\
+ Sp_{B} \sum_{\left(s\right)} \left(- \frac{i}{\hbar}\right)^{s}\int_{t_0}^{t} dt_{s} \int_{t_0}^{t_s} dt_{s-1} \dots
\nonumber \\
\dots \int_{t_0}^{t_{2}} dt_1 
\left[\hat{V}_{AB}\left(t_s\right), 
\left[\hat{V}_{AB}\left(t_{s - 1}\right),
\dotsc
\left[\hat{V}_{AB}\left(t_1\right),
\hat{\rho}_{A}\left(t_0\right)
\otimes
\hat{\rho}_{B}\left(t_0\right)
\right]
\dots
\right]
\right].
\label{eqCh2_rho_sequance}
\end{eqnarray}
Usually, only the first few terms
of the expansion are considered. For example, the first two terms of the expansion are:
\begin{eqnarray}
\hat{\rho}_{A}\left(t\right) = 
\hat{\rho}_{A}\left(t_0\right) - \frac{i}{\hbar} Sp_B \left\{
\int_{t_0}^{t} 
\left[\hat{V}_{AB}\left(t_1\right), \hat{\rho}_{A}\left(t_0\right)
\otimes
\hat{\rho}_{B}\left(t_0\right)
\right]d t_1
\right\} -
\nonumber \\
-
\frac{1}{\hbar^2}
Sp_B \left\{
\int_{t_0}^{t}d t_2 
\int_{t_0}^{t_2} d t_1
\left[\hat{V}_{AB}\left(t_2\right), 
\left[\hat{V}_{AB}\left(t_1\right), 
\hat{\rho}_{A}\left(t_0\right)
\otimes
\hat{\rho}_{B}\left(t_0\right)
\right]
\right]
\right\}.
\nonumber
\end{eqnarray}