%% -*- coding:utf-8 -*- 
\subsection{Finding the Period of Functions Using Quantum Fourier Transform}

To determine the period of the function \eqref{eqPart4QuantCompShorClassPart}, the scheme shown in \autoref{figQuantCompQuantPeriodFinding} is used.

\input part4/quantcomp/figquantperiodfinding.tex

The first element is the Hadamard transform for $n$ qubits, \rindex{Hadamard Transform} which prepares the initial state as follows:
\begin{equation}
\ket{in} = \frac{1}{\sqrt{2^n}}\sum_{x=0}^{2^n - 1} \ket{x} \otimes \ket{0}.
\nonumber
\end{equation}

After the element computing the function $\hat{U}_f$, we have for the state
\begin{equation}
\hat{U}_f\ket{in} = \frac{1}{\sqrt{2^n}}\sum_{x=0}^{2^n - 1} \ket{x} \otimes \left|f\left(x\right)\right>.
\nonumber
\end{equation}

\input part4/quantcomp/figshorquant.tex

After measuring the function value, only those elements for which the function value equals the measured value will remain in the list of coordinates. As a result, the input to the Fourier transformation measurement element is a state of the form
\begin{equation}
\ket{in'} = \sum_{x'} \ket{x'},
\nonumber
\end{equation}
where all non-zero elements have the same amplitude and follow with a period equal to the period of the investigated function. The initial value will be shifted depending on the experiment (the shift will vary in different experiments). By lemma \ref{lemmaAddFourierDiscretFourierShiftTime}, the Fourier image will be the same for different function measurements.

Further, by lemma \ref{lemmaAddFourierDiscretFourierPeriod}, (on periodicity) (see also comment \ref{rem:dsp:fourier:periodprop}), it follows that the most probable readings (probability maxima) follow with a period related to the original period of the function. Thus, as a result of several experiments, the desired function period can be found with the required probability level (see \autoref{picPart4QuantCompShorQuantPart}).

\begin{example}
\emph{Finding the period of the function $f\left(x\right) = 2^x \mod 21$}
\label{exPart4QuantCompShorQuantPeriodFinding}
As an example, consider the task of finding the period of the function $f\left(x, a\right) = a^x \mod{N}$ with $a=2$, $N=21$; see \autoref{picPart4QuantCompShorQuantPart}.

The number of samples $M$ must be a power of two. In our example, we choose $M = 2^6 = 64$ as the number of samples. Thus, 6 qubits are needed for our example.

The initial state after the Hadamard transform is as follows:
\begin{equation}
\ket{in} = \frac{1}{8}\sum_{x = 0}^{63}\ket{x} \otimes \ket{0},
\nonumber
\end{equation}
where $\ket{x}$ represents the tensor product \rindex{Tensor Product} of 6 qubits that encode the binary representation of the function argument. For example, for $x=5_{10}=000101_2$ we have
\[
\ket{x} = \ket{0}\otimes \ket{0}\otimes \ket{0}\otimes \ket{1}\otimes \ket{0}\otimes \ket{1}
\]

After computing the function, we have a state of the form (see the top plot in \autoref{picPart4QuantCompShorQuantPart})
\begin{eqnarray}
\hat{U}_f\ket{in} = \frac{1}{8}\sum_{x = 0}^{63}\ket{x} \otimes \left|f\left(x\right)\right> = 
\nonumber \\
=
\frac{1}{8}
\left(
\ket{0}\otimes\ket{2} + 
\ket{1}\otimes\ket{4} + 
\ket{2}\otimes\ket{8} + \dots +
\right.
\nonumber \\
\left.
+
\ket{62}\otimes\ket{8} +
\ket{63}\otimes\ket{16}
\right).
\label{eqPart4QuantCompShorPFExample1}
\end{eqnarray}

If the result of the function measurement was $1$, then from the entire sum \eqref{eqPart4QuantCompShorPFExample1}, the terms for which the function value equals $1$ will remain (see the middle plot in \autoref{picPart4QuantCompShorQuantPart}):
\begin{equation}
\ket{in'} = \frac{1}{\sqrt{10}}\left( 
\ket{5}\otimes\ket{1} +
\ket{11}\otimes\ket{1} +
\ket{17}\otimes\ket{1} +
\dots +
\ket{60}\otimes\ket{1}
\right).
\label{eqPart4QuantCompShorPFExample2}
\end{equation} 
Expression \eqref{eqPart4QuantCompShorPFExample2} contains 10 terms of equal amplitude, so the normalization factor is $\frac{1}{\sqrt{10}}$.

The Fourier transform for the sequence \eqref{eqPart4QuantCompShorPFExample2} is shown in the bottom plot \autoref{picPart4QuantCompShorQuantPart}. The most probable values of the Fourier transform measurement result will correspond to local maxima that repeat with a period of $\frac{M}{r}\approx10.67$ from which the period of the desired function $r=6$ can be found.

\end{example}