%% -*- coding:utf-8 -*- 
\section{Quantization of the Free Electromagnetic Field}
\label{Ch1_quantumemf}
\rindex{quantization!free electromagnetic field}
Using the analogy noted earlier, one can quantize
the electromagnetic field similarly to how a simple
mechanical harmonic oscillator is quantized.

Upon quantization, $q_n$ and $p_n$ become operators $\hat{q}_n$,
$\hat{p}_n$, satisfying the same commutation relations as
coordinate and momentum:
\begin{equation}
\left[\hat{q}_n, \hat{p}_n\right] = \hat{q}_n\hat{p}_n -
\hat{p}_n \hat{q}_n = i\hbar
\label{eqCh1_comut}
\end{equation}
where $\hat{q}_n$ and $\hat{p}_n$ are Hermitian (self-adjoint)
operators. In the Schrödinger representation they are time-independent, in the Heisenberg representation they are time-dependent. It is convenient to introduce new operators through these, which are not self-adjoint but are conjugates of each other:
\begin{equation}
\hat{a}_n = \frac{1}{\sqrt{2 \hbar \omega_n}}
\left( \omega_n \hat{q}_n + i \hat{p}_n\right),
\quad
\hat{a}_n^{\dag} = \frac{1}{\sqrt{2 \hbar \omega_n}}
\left( \omega_n \hat{q}_n - i \hat{p}_n\right),
\label{eqCh1_aadef}
\end{equation}
The inverse relations yield
\begin{equation}
\hat{q}_n = \sqrt{\frac{\hbar}{2 \omega_n}}
\left(\hat{a}_n + \hat{a}_n^{\dag}\right),
\quad
\hat{p}_n = i \sqrt{\frac{\hbar \omega_n}{2}}
\left(\hat{a}_n^{\dag} - \hat{a}_n\right),
\label{eqCh1_qpdef}
\end{equation}

Using expressions \eqref{eqCh1_comut}, 
\eqref{eqCh1_aadef}, \eqref{eqCh1_qpdef} one can obtain commutation
relations for the operators $\hat{a}_n$ and $\hat{a}_n^{\dag}$:
\begin{eqnarray}
\left[\hat{a}_n, \hat{a}_n^{\dag}\right] = 
\frac{1}{2 \hbar \omega_n}
\left( \omega_n \hat{q}_n + i \hat{p}_n\right) 
\left( \omega_n \hat{q}_n - i \hat{p}_n\right) - 
\nonumber \\
- \frac{1}{2 \hbar \omega_n}
\left( \omega_n \hat{q}_n - i \hat{p}_n\right) 
\left( \omega_n \hat{q}_n + i \hat{p}_n\right) =
\nonumber \\
= \frac{1}{2 \hbar \omega_n}
\left( - 2 i \omega_n 
\left(\hat{q}_n \hat{p}_n - \hat{p}_n \hat{q}_n\right)\right) = 1.
\end{eqnarray}
Thus,
\begin{equation}
\left[\hat{a}_n, \hat{a}_n^{\dag}\right] = 1.
\label{eqCh1_aacomutation}
\end{equation}
Moreover, it is obvious that
\begin{equation}
\left[\hat{a}_n, \hat{a}_n\right] = 0,
\quad
\left[\hat{a}_n^{\dag}, \hat{a}_n^{\dag}\right] = 0.
\end{equation}
\rindex{Hamiltonian}
The Hamiltonian, expressed through the new operators, takes the form
\begin{eqnarray}
\hat{\mathcal{H}_n} = 
\frac{1}{2}\left(\omega_n^2 \hat{q}_n^2 + \hat{p}_n^2\right) = 
\nonumber \\
= \frac{1}{2}\frac{\omega_n \hbar}{2}
\left(
\left(\hat{a}_n + \hat{a}_n^{\dag} \right)
\left(\hat{a}_n + \hat{a}_n^{\dag} \right)
-
\left(\hat{a}_n^{\dag} - \hat{a}_n \right)
\left(\hat{a}_n^{\dag} - \hat{a}_n \right)
\right) =
\nonumber \\
= \frac{\omega_n \hbar}{2} 
\left( \hat{a}_n \hat{a}_n^{\dag} + \hat{a}_n^{\dag} \hat{a}_n\right) =
\frac{\omega_n \hbar}{2} 
\left(1 + \hat{a}_n^{\dag} \hat{a}_n + \hat{a}_n^{\dag} \hat{a}_n\right) =
\nonumber \\
= \omega_n \hbar 
\left(\hat{a}_n^{\dag} \hat{a}_n + \frac{1}{2}\right)
\label{eqCh1_quant_stoyachie_volny}
\end{eqnarray}

Here the relations between $\hat{q}_n$, $\hat{p}_n$ 
and $\hat{a}_n^{\dag}$, $\hat{a}_n$ \eqref{eqCh1_aadef} are used, as well as
the commutation relation \eqref{eqCh1_aacomutation}.

The energy $\frac{1}{2}\omega_n \hbar$ corresponds to zero-point
oscillations. The total zero-point energy
\[
\frac{1}{2}\sum_{(n)}\omega_n \hbar \to \infty,
\] 
since the number of modes is infinite. This does not lead to 
significant problems because we are only interested in energy differences,
and the constant term can be discarded. Then the Hamiltonian
takes the form
\begin{equation}
\hat{\mathcal{H}} = \sum_{(n)}\hat{\mathcal{H}_n} =
\sum_{(n)}\hbar \omega_n \hat{a}_n^{\dag}\hat{a}_n
\end{equation}

The electric and magnetic field operators can be expressed as
\begin{eqnarray}
\hat{\vec{E}} = \sum_{(n)}\frac{\hat{q}_n
  \omega_n}{\sqrt{\varepsilon_0}} \vec{E}_n\left(r\right) = 
\sum_{(n)}\sqrt{\frac{\hbar \omega_n}{2 \varepsilon_0}}
\left(\hat{a}_n^{\dag} + \hat{a}_n \right)
\vec{E}_n\left(r\right),
\nonumber \\
\hat{\vec{H}} = \sum_{(n)}\frac{\hat{p}_n}
{\sqrt{\mu_0}} \vec{H}_n\left(r\right) = 
i \sum_{(n)}\sqrt{\frac{\hbar \omega_n}{2 \mu_0}}
\left(\hat{a}_n^{\dag} - \hat{a}_n \right)
\vec{H}_n\left(r\right).
\end{eqnarray}

Let us consider a simple example. In laser theory one considers a resonator
formed by two parallel mirrors. Often, an approximate field representation
is used, assuming the field depends only on the longitudinal coordinate (\autoref{figCh1_Res}). In
this approximation, the normalized eigenfunction (mode) can be
represented as:
\begin{equation}
E_{nx} = \sqrt{\frac{2}{V}} \sin k_n z,
\quad
H_{ny} = \sqrt{\frac{2}{V}} \cos k_n z,
\quad
k_n = \frac{\pi n}{L},
\quad
\omega_n = \frac{c \pi n}{L},
\end{equation}
where  $V = SL$,  $L$ is the resonator length,  $S$ is the cross-section of the light beam.

\input ./part1/quantel/fig1.tex

The operator of the mode's electric field in this case will be:
\begin{eqnarray}
\hat{E}_{nx} = 
\sqrt{\frac{\hbar \omega_n}{2 \varepsilon_0}}
\left(\hat{a}_n^{\dag} + \hat{a}_n \right)
\sqrt{\frac{2}{V}} \sin k_n z = 
E_1 \left(\hat{a}_n^{\dag} + \hat{a}_n \right) \sin k_n z, 
\nonumber \\
\hat{H}_{ny} = i E_1 \sqrt{\frac{\varepsilon_0}{\mu_0}}
\left(\hat{a}_n^{\dag} - \hat{a}_n \right) \cos k_n z,
\label{eqCh1_EH_simple}
\end{eqnarray}
where
$E_1 = \sqrt{\frac{\hbar \omega}{\varepsilon_0 V}}$ is the
electric field corresponding to one photon (quantum) in
the mode. 
