%% -*- coding:utf-8 -*-
\section{Boolean Logic. CNF Boolean Formulas}

\begin{definition}[Conjunction (logical ``AND'')]
For the logical operation ``AND'' ($a \& b$), the following truth table
\ref{tblAddAlgoAND} holds.
\begin{table}
\centering
\begin{tabular}{|c|c|c|}
\hline
$a$ & $b$ & $a \& b$ \\ \hline
0  & 0 & 0 \\
0  & 1 & 0 \\
1  & 0 & 0 \\
1  & 1 & 1 \\ \hline
\end{tabular}
\caption{Conjunction $a \& b$}
\label{tblAddAlgoAND}
\end{table}
\end{definition}

\begin{definition}[Disjunction (logical ``OR'')]
For the logical operation ``OR'' ($a \| b$), the following truth table
\ref{tblAddAlgoOR} holds.
\begin{table}
\centering
\begin{tabular}{|c|c|c|}
\hline
$a$ & $b$ & $a \| b$ \\ \hline
0  & 0 & 0 \\
0  & 1 & 1 \\
1  & 0 & 1 \\
1  & 1 & 1 \\ \hline
\end{tabular}
\caption{Disjunction $a \| b$}
\label{tblAddAlgoOR}
\end{table}
\end{definition}

\begin{definition}[Boolean Formula]
A Boolean formula is defined as a set of Boolean literals
combined by logical operations.
\end{definition}

\begin{definition}[CNF - Conjunctive Normal Form (SAT)]
A conjunctive normal form in Boolean logic is called a Boolean
formula that has the form of a conjunction of disjunctions of literals.
\end{definition}

\begin{example}
\emph{SAT Problem}
\label{exAddAlgoSAT}
The SAT problem is the problem of satisfiability of CNF. Suppose the following
Boolean formula in CNF form is given:
\begin{equation}
\left(x_1 \| x_2 \| \bar{x_3} \right) \& \left(x_1 \| \bar{x_2} \| x_3 \right)
\nonumber
\end{equation}
It is satisfiable; it suffices to assign $x_1 = 1$, and the values of
$x_{2,3}$ can be arbitrary.

At the same time, the following formula is unsatisfiable:
\begin{equation}
\left(x_1 \| x_2 \| \bar{x_3} \right) \& 
\left(\bar{x_1} \| \bar{x_2} \| x_3 \right).
\nonumber
\end{equation}

\end{example}

\begin{theorem}[Reduction to CNF]
Any Boolean formula can be reduced to CNF. 
\end{theorem}

\begin{proof}
For the proof, one can use: the law of double negation, De Morgan's law,
and distributivity.
\end{proof}

\begin{theorem}[Reduction to 3-CNF (3-SAT)]
Any Boolean formula in conjunctive normal form can be
reduced to 3-CNF.  
\end{theorem}

\begin{proof}
TBD
\end{proof}